---
title: "Analyzing Sakura Flowering Trends Under Climate Change"
subtitle: "Predicting Bloom Dates Using Historical Records and Temperature Data"
author: 
  - Mariko Lee
thanks: "Code and data are available at: [https://github.com/leemarik/sakura-flowering-trends.git](https://github.com/leemarik/sakura-flowering-trends.git)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```
```{r}
#| include: false
#| warning: false
#| message: false

#### Workspace setup ####
library(broom.mixed)
library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
library(here)
library(kableExtra)
library(arrow)
library(rstanarm)

#### Load necessary data and models ####
clean_sakura_data <- read_parquet("../data/02-analysis_data/clean_sakura_data.parquet")
lm_mode <- readRDS("../models/lm_model.rds")
gam_model <- readRDS("../models/gam_model.rds")

```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}


## Overview


Overview text

## Measurement
	
Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.



Talk more about it.

Talk way more about it. 

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.



\newpage

# Model {#sec-model}
To determine the impact of temperature on sakura flowering dates over centuries and develop predictive models for future trends under different climate scenarios, I used a linear regression model (LM) as a baseline and a generalized additive model (GAM) for capturing non-linear relationships in the data. 

## Linear Regression Model 
The linear regression model I will be using is:

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

## Generalized Additive Model (GAM)

Where:


### Model justification



# Results


# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics


\newpage


# References


